
Properties were presented in the form ``This one X. Who else do you think X''.

\textbf{Properties specific to Flurps (animals)}

\begin{itemize} 
  \singlespacing
    \item This one breathes. 
    \item This one can have babies. 
    \item This one has a heart inside. 
    \item This one has a mummy. 
    \item This one has bones inside. 
    \item This one needs water. 
    \item This one can climb trees. 
    \item This one tries to stay warm. 
\end{itemize}

\textbf{Properties specific to Floobits (robots)}
\begin{itemize}
\singlespacing
    \item This one can be turned off. 
    \item This one can break. 
    \item This one has batteries inside. 
    \item This one has wires inside. 
    \item This one was made by people. 
    \item This one was sold in a store. 
    \item This one is cold touch. 
    \item This one doesn't sleep. 
\end{itemize}

\textbf{Generic properties}
\begin{itemize}
\singlespacing
    \item This one can make a zevy sound. 
    \item This one has a very sticky toma. 
    \item This one has zimmer inside. 
    \item This one is used for derriping. 
    \item This one needs tiddles to make it move. 
    \item This one uses danner. 
    \item This one can help yippets. 
    \item This one goes outside in the winter. 
    \item This one has a part inside called a cece. 
    \item This one has blickets inside of it. 
    \item This one has grumpets that make it strong. 
    \item This one is good for kertling. 
    \item This one is found on farms. 
    \item This one can be very old. 
    \item This one feels yinty. 
    \item This one lacks ombelots. 
\end{itemize}


